---
title: |
  ProsperOK Metrics Report<br>for Tulsa County Prison Population,<br> Q4 2025
subtitle: |
  November 2025 update on Tulsa County Prison Population and Sentences
format:
  ojo-report-template-html:
    theme: default
    mainfont: "Lora"
    header-includes: |
      <link rel="preconnect" href="https://fonts.googleapis.com">
      <link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
      <link href="https://fonts.googleapis.com/css2?family=Lato:wght@100;300;400;900&display=swap" rel="stylesheet">
      <link href="https://fonts.googleapis.com/css2?family=Lato:wght@100;300;400;900&family=Lora:ital,wght@0,400;0,500;0,600;0,700;1,400;1,500;1,600;1,700&display=swap" rel="stylesheet">
# Formatting options -----------------------------------------------------------
logo: "www/okpi-square.png" # Other options are in the /www/ directory
number-sections: true
smooth-scroll: true
title-block-banner: "#2f507f" # The background color for the header banner
title-block-banner-color: "white" # The text color for the header banner
fontcolor: "black" # Default text color for the body
linkcolor: "black" # Default link color for the body
toc: true # This should be 'false' if your end goal is a PDF
knitr:
  opts_chunk:
    fig.align: center
    echo: false
    message: false
    warning: false
    fig-width: 7
    fig-height: 5
    out-width: '100%'
# Document info ----------------------------------------------------------------
author: Anthony Flores
date: last-modified
description: |
    This is a status update report using data from the Tulsa County Prison. It was prepared and analyzed by the Oklahoma Policy Institute.
embed-resources: false
# abstract-title: Publication Note
# abstract: 'DRAFT REPORT -- NOT FOR PUBLICATION'
---

```{r}

# To render to PDF after you've rendered your .html document:
# pagedown::chrome_print("./report.html", "./report.pdf", format = "pdf")

# We'll build this into the YAML header so you don't have to take this extra step
# once the {pagedown} people get around to adding quarto support.

```

```{r setup, include=FALSE}
library(here)
library(ojodb)
library(ojothemes)
library(janitor)
library(gt)
library(gtExtras)
library(tidyverse)
library(nanoparquet)
library(ggrepel)

ojothemes::ojo_set_theme(theme = "okpi")
```

# Overview

This is an overview report to provide some basic numbers about the current
population and recent bookings at the Oklahoma County Detention Center (OCDC).

It includes aggregations for each in total and by gender:

* Prison Sentences from Tulsa County
* Prison Releases from those sentenced in Tulsa County
* Prison Population by Gender  from Tulsa County
* Oklahoma County comparison to Tulsa County vs all other counties

{{< pagebreak >}}

# Prison Sentences

# Prison Releases

# Population by Gender

# Oklahoma County comparison to Tulsa County vs all other counties
