---
title: |
  ProsperOK Metrics Report<br>for Tulsa County Jail,<br> Q4 2025
subtitle: |
  November 2025 update on Tulsa County Jail Population and Bookings
format:
  ojo-report-template-html:
    theme: default
    mainfont: "Lora"
    header-includes: |
      <link rel="preconnect" href="https://fonts.googleapis.com">
      <link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
      <link href="https://fonts.googleapis.com/css2?family=Lato:wght@100;300;400;900&display=swap" rel="stylesheet">
      <link href="https://fonts.googleapis.com/css2?family=Lato:wght@100;300;400;900&family=Lora:ital,wght@0,400;0,500;0,600;0,700;1,400;1,500;1,600;1,700&display=swap" rel="stylesheet">
# Formatting options -----------------------------------------------------------
logo: "www/okpi-square.png" # Other options are in the /www/ directory
number-sections: true
smooth-scroll: true
title-block-banner: "#2f507f" # The background color for the header banner
title-block-banner-color: "white" # The text color for the header banner
fontcolor: "black" # Default text color for the body
linkcolor: "black" # Default link color for the body
toc: true # This should be 'false' if your end goal is a PDF
knitr:
  opts_chunk:
    fig.align: center
    echo: false
    message: false
    warning: false
    fig-width: 7
    fig-height: 5
    out-width: '100%'
# Document info ----------------------------------------------------------------
author: Anthony Flores
date: last-modified
description: |
    This is a status update report using data from the Tulsa County Jail. It was prepared and analyzed by the Oklahoma Policy Institute.
embed-resources: false
params:
  public: false
# abstract-title: Publication Note
# abstract: 'DRAFT REPORT -- NOT FOR PUBLICATION'
---

```{r}

# To render to PDF after you've rendered your .html document:
# pagedown::chrome_print("./report.html", "./report.pdf", format = "pdf")

# We'll build this into the YAML header so you don't have to take this extra step
# once the {pagedown} people get around to adding quarto support.

```

```{r setup, include=FALSE}
library(here)
library(ojodb)
library(ojothemes)
library(janitor)
library(gt)
library(gtExtras)
library(tidyverse)
library(nanoparquet)
library(ggrepel)

ojothemes::ojo_set_theme(theme = "okpi")

analysis_path <- Sys.getenv("PROSPER_JAIL_ANALYSIS_RDS")
figures_path <- Sys.getenv("PROSPER_JAIL_FIGURES_RDS")

jail_analysis <- if (nzchar(analysis_path) && file.exists(analysis_path)) readRDS(analysis_path) else NULL
jail_figures <- if (nzchar(figures_path) && file.exists(figures_path)) readRDS(figures_path) else NULL

fmt_pct <- function(x) scales::percent(x, accuracy = 0.1)
fmt_num <- function(x) scales::comma(x, accuracy = 1)
```

# Overview


# Data Challenges in 2025
In 2025, Tulsa County Jail (the David L. Moss Criminal Justice Center),
migrated its Jail Management System to a new vendor, 365Labs Public Safety Platform,
which impacted data availability and reporting capabilities for much of the year.
The jail's internal reporting tools were not fully functional for some time after
the migration, and prior web scrapers became non-fuctional after the new
community data portal was launched.

Given these unique challenges, this report attempts to provide an update on key
metrics for the jail population and bookings by using a variety of data sources to
triangulate estimates where possible.

Over the coming months, we will continue work with Tulsa County jail
administrators to improve data access and reporting capabilities to ensure
accurate and timely information is available moving forward. We will also be
able to use that data to validate and refine the estimates presented in this report.


```{r overview-metrics}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)
if (!is.null(jail_analysis)) {
  jail_analysis$key_metrics |>
    dplyr::mutate(
      value = fmt_num(value),
      yoy_change = dplyr::case_when(
        is.na(yoy_change) ~ "—",
        TRUE ~ fmt_pct(yoy_change)
      )
    ) |>
    dplyr::rename(`Metric` = metric, `Value` = value, `YoY Change` = yoy_change) |>
    gt::gt() |>
    gtExtras::gt_theme_538()
} else {
  "Pipeline analysis has not been loaded yet."
}
```

{{< pagebreak >}}

# Jail Bookings

```{r}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)
if (!is.null(jail_figures$booking_trend)) {
  jail_figures$booking_trend
}
```

```{r jail-bookings-table}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)
if (!is.null(jail_analysis)) {
  latest <- jail_analysis$latest_month |>
    dplyr::mutate(
      `Latest Month` = format(booking_month, "%B %Y"),
      Bookings = fmt_num(bookings)
    ) |>
    dplyr::select(Source = source, `Latest Month`, Bookings)

  annual <- jail_analysis$booking_year_totals |>
    dplyr::arrange(dplyr::desc(booking_year)) |>
    dplyr::mutate(
      `Year` = booking_year,
      Bookings = fmt_num(total_bookings)
    ) |>
    dplyr::select(Source = source, Year, Bookings)

  dplyr::bind_rows(
    dplyr::mutate(latest, Table = "Latest Month"),
    dplyr::mutate(annual, Table = "Year-to-date")
  ) |>
    gt::gt(rowname_col = "Source", groupname_col = "Table") |>
    gtExtras::gt_theme_538()
}
```

```{r jail-releases-heading, results='asis', echo=FALSE}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)
if (!params$public) {
  cat("# Jail Releases\n\n")
}
```

```{r jail-releases}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)
#| results: "hide"

release_counts <- tibble::tibble()
release_shares <- tibble::tibble()
if (!is.null(jail_analysis)) {
  release_counts <- jail_analysis$release_counts |>
    dplyr::filter(group %in% c("All", "Male", "Female")) |>
    dplyr::mutate(Value = fmt_num(value)) |>
    dplyr::select(Disposition = category, Group = group, Value)

  release_shares <- jail_analysis$release_shares |>
    dplyr::mutate(Value = fmt_pct(value)) |>
    dplyr::select(Disposition = category, Group = group, Value)

  list(
    release_counts = release_counts,
    release_shares = release_shares
  )
}
```

```{r jail-release-counts}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)

if (nrow(release_counts) > 0) {
  release_counts |>
    gt::gt() |>
    gtExtras::gt_theme_538()
} else {
  "Release counts are not available for this run."
}
```

```{r jail-release-shares}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)

if (nrow(release_shares) > 0) {
  release_shares |>
    gt::gt() |>
    gtExtras::gt_theme_538()
} else {
  "Release percentages are not available for this run."
}
```

```{r jail-comparison-heading, results='asis', echo=FALSE}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)
if (!params$public) {
  cat("# Tulsa County comparison to Oklahoma County\n\n")
}
```

```{r jail-comparison}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)

if (!is.null(jail_analysis)) {
  jail_analysis$brek_report |>
    dplyr::filter(
      metric_family %in% c("adp", "bookings"),
      dimension == "Overall",
      category == "Total",
      county %in% c("Tulsa County", "Oklahoma County")
    ) |>
    dplyr::mutate(
      Metric = dplyr::case_when(
        metric_family == "adp" ~ "Average Daily Population",
        TRUE ~ "Annual Bookings"
      ),
      Value = fmt_num(value),
      `YoY Change` = fmt_pct(as.numeric(yoy_change))
    ) |>
    dplyr::select(Metric, County = county, Value, `YoY Change`) |>
    gt::gt() |>
    gtExtras::gt_theme_538()
}
```

# Average Daily Population

```{r jail-adp}
#| include: !expr isFALSE(params$public)
#| eval: !expr isFALSE(params$public)
if (!is.null(jail_figures$adp_comparison)) {
  jail_figures$adp_comparison
}
```
