---
title: HB 1795 Impact Estimate
subtitle: Open Justice Oklahoma
format:
  ojo-report-template-html: default
  pdf:
    pdf-engine: xelatex
author: Polina Rozhkova
date: last-modified
logo: "www/ojo-logo-white.png"
number-sections: true
smooth-scroll: true
title-block-banner: "black"
title-block-banner-color: "white"
knitr:
  opts_chunk:
    fig.align: center
    echo: FALSE
    message: FALSE
    warning: FALSE
execute: 
  freeze: auto
description: Estimating the number of Oklahomans impacted by the amended provision concerning driver license revocations for individuals with misdemeanor drug charges. Analysis conducted at the request of Oklahomans for Criminal Justice Reform (OCJR) and ProsperOK.
abstract-title: "Abstract"
abstract: Between 9,187 and 16,116 Oklahomans with misdemeanor drug charges were impacted by provisions in HB 1795 between January 1, 2022 and December 1, 2023.
---

```{r}
library(readr)
library(stringr)
library(readxl)
library(tidycensus)
library(tidyverse)
library(here)
library(janitor)
library(glue)
library(ojoverse)
library(lubridate)
library(gt)
library(downloadthis)
library(kableExtra)
library(gtExtras)
```

## Background

Oklahoma’s [House Bill 1795](http://webserver1.lsb.state.ok.us/cf_pdf/2021-22%20ENR/hB/HB1795%20ENR.PDF) went into effect November 1, 2021 (Q2 FY 2022). This new law reduces the number of offenses that can lead to a license revocation, although certain circumstances may still result in revocation. Additionally, HB 1795 sought to enhance the accessibility and affordability of provisional licenses. The legislation imposed restrictions on license suspensions resulting from "failure to pay" and incorporated a significant provision reducing the minimum payment plan for outstanding fines and fees. Previously set at \$25 per month, HB 1795 lowers the minimum monthly payment to \$5. The aforementioned fees encompass various charges such as provisional license fees, warrant fees, court costs or fees, and driver license reinstatement fees.

## Data and Methodology

Open Justice Oklahoma maintains a database of administrative court records which includes information on all criminal misdemeanors and felonies filed in Oklahoma beginning in 2001. Case information is systematically collected from publicly available data hosted on the [Oklahoma State Court Network (OSCN)](https://www.oscn.net/v4/) website. This analysis uses Oklahoma court records beginning January 1, 2022 (Q3 FY 2022) to estimate the impact of HB 1795 during a period of full implementation. We explore two provisions of the bill in order to estimate lower and upper bounds for the number of individuals with misdemeanor drug charges that were impacted by the bill’s passing.

To estimate the lower bound, we focus on the provision concerned with modifying the types of offenses requiring immediate revocation of driving privileges (Section 6-205). Using the Department of Safety [violation codes](https://oklahoma.gov/content/dam/ok/en/dps/VCB%20February%202022.pdf) we pull court records for individuals who are charged with possession of a controlled dangerous substance as defined by Oklahoma’s [Uniform Controlled Dangerous Substances Act](http://www.oklegislature.gov/cf_pdf/2003-04%20INT/hb/HB2166%20int.pdf) while driving a motor vehicle. Court records also include drug violation codes that are similar but do not specify use of a motor vehicle. These violation codes provide a minimum estimate of individuals whose drug charges would have resulted in immediate driver license suspension or revocation prior to the enactment of HB 1795.

Prior to the bill’s passing, individuals could have their license suspended or revoked for the failure to pay court fines or fees. Some of these fines and fees are imposed on individuals who are charged with misdemeanor drug offenses. Following this logic, we account for any individuals who may be subject to fines and/or fees due to a misdemeanor drug possession charge. We combine the previously calculated lower bound and adjust for duplicate entries (individuals that may appear in both methods) for an upper bound estimate. 

```{r}
drugs <- "(?i)CDS|C\\.D\\.S|DRUG|OXY|HUFF|AMPHET|ZOLOL|ZOLAM|HYDROC|CODEIN|PRECURS|XANAX|MORPH|METERDI|ZEPAM|LORAZ|VALIU|EPHED|SUB|COCA|PSEUDO| CS|CS | CD|CD |PRESCRIP|NARC|METH|C\\.D\\.|HEROIN|ANHYD|AMMONIA|OPIUM|LORTAB|MARIJUANA|MARIHUANA|MJ"
drug_exclude <- "(?i)DOMESTIC|ASSAULT|FIREARM|DISTRIBUTE|INTENT|MANUFACTURE|DISPENSE|TUO|alcohol|acohol|alchol"
vehicle_related <- "VEHICLE"

# DPS violation codes saved in "docs" along with bill
dps_codes <- "(?i)DI1|DI1M|DI2D|DI2M|DI2DM|DRI|DU2II|DU2IV|DU4|DU4II|DU8I|DU8II|DU9|DU9I|DU9II|DU9IV|dui drugs misdemeanor"
dps_exclude <- "(?i)alcohol|acohol|alchol|under 21|suspended|revoked|valid|twenty-one"
dps_related <- "(DPS|DEPARTMENT OF PUBLIC SAFETY|D\\.P\\.S)"

min_desc_dl <- "(SUSPENSION|SUSPEND)"
min_desc_exclude <- "(WITHDRAW|ERROR|RETURN|UNABLE|REVOKE|LIFT|RECALL|NOT PROCESSED|PENDING|SUSPENSION RELEASE|DEFENDANT APPEARS)"

# Set to reporting period
# Replace with fiscal quarter using ojo_fiscal_year
cumulative_period_start <- ymd("2022-01-01")
reporting_period_start <- ymd("2023-01-01")
reporting_period_end <- ymd("2023-12-01")
```


```{r, cache=TRUE}

ojodb <- ojo_connect()

all_cases <- ojo_tbl("case", .con = ojodb) |>
  filter(
    case_type == "CM",
    date_filed >= cumulative_period_start,
    date_filed <= reporting_period_end
  ) |>
  select(
    id,
    case_type,
    district,
    date_filed,
    counts,
    open_counts
  ) |>
  mutate(
    count = unnest(counts),
    open_count = unnest(open_counts)
  ) |>
  left_join(
    ojo_tbl("count", .con = ojodb),
    by = c("count" = "id"),
    suffix = c("", ".count")
  ) |>
  mutate(
    count_as_filed = if_else(
      is.na(open_count),
      count_as_filed,
      open_count
    )
  ) |>
  bind_rows(
    ojo_tbl("case", schema = "oscn", .con = ojodb) |>
      left_join(
        ojo_tbl("count", schema = "oscn", .con = ojodb),
        by = c("id" = "case_id")
      )
  ) |>
  collect()

# Defendants for all cases
total_persons <- ojo_tbl("party", .con = ojodb) |>
  bind_rows(
    ojo_tbl("party", schema = "oscn", .con = ojodb)
  )
  right_join(
    all_cases |>
      distinct(id),
    by = c("case_id" = "id"),
    copy = TRUE
  ) |>
  filter(role == "Defendant") |>
  distinct(oscn_id) |>
  count() |>
  collect()

# Pulling every case with at least one misdemeanor drug charge
# (excluding distribution/intent to dispense/etc.)
case_misdemeanor_drug <- 
  all_cases |>
  filter(
    str_detect(count_as_filed, drugs)
  ) |>
  filter(
    !str_detect(count_as_filed, drug_exclude)
  )

# Defendants in cases above:
drug_misdemeanor_persons <- ojo_tbl("party", .con = ojodb) |>
  bind_rows(ojo_tbl("party", schema = "oscn", .con = ojodb)) |>
  right_join(
    case_misdemeanor_drug |>
      distinct(id),
    by = c("case_id" = "id"),
    copy = TRUE
  ) |>
  filter(role == "Defendant") |>
  distinct(oscn_id) |>
  count() |>
  collect()

# A conservative estimate of the number of impacted individuals can be made by
# filtering misdemeanors by DPS violation code and related codes.
# For the most part, this excludes drug offenses occurring outside of motor vehicle
case_dps_violation <- 
  all_cases |>
  filter(str_detect(count_as_filed, dps_codes),
         !str_detect(count_as_filed, dps_exclude))

# This calculates number of defendants:
dps_violation_persons <- ojo_tbl("party", .con = ojodb) |>
  bind_rows(ojo_tbl("party", schema = "oscn", .con = ojodb)) |>
  right_join(
    case_dps_violation |>
      distinct(id),
    by = c("case_id" = "id"),
    copy = TRUE
  ) |>
  filter(role == "Defendant") |>
  distinct(oscn_id) |> 
  count() |>
  collect()

# Upper Bound including people from DPS OR misdemeanor method:
both_methods_persons <- ojo_tbl("party", .con = ojodb) |>
  bind_rows(
    ojo_tbl("party", .con = ojodb)
  ) |>
  right_join(
    rbind(
      case_dps_violation |>
        select(id, date_filed),
      case_misdemeanor_drug |>
        select(id, date_filed)
    ) |>
      distinct(),
    by = c("case_id" = "id"),
    copy = TRUE
  ) |>
  filter(role == "Defendant") |>
  distinct(oscn_id, .keep_all = TRUE) |> 
  #distinct(oscn_id) |>
  #count() |>
  collect()

```

## Results

```{r}
both_methods_persons |> 
  distinct(oscn_id) |>
  count() |> 
  bind_cols(
  dps_violation_persons,
  #drug_misdemeanor_persons,
  #both_methods_persons) |>
  ) |> 
  gt() |> 
  # cols_label(contains("1") ~ "Lower Bound", 
  #            contains("2") ~ "Upper Bound") |>
  cols_label(contains("2") ~ "Lower Bound", 
             contains("1") ~ "Upper Bound")|>
  tab_header(
    title = "Estimated Impact for Oklahomans with Misdemeanor Drug Charges",
    subtitle = glue::glue("{cumulative_period_start} to {reporting_period_end}")) |> 
  tab_style(
    style = cell_text(color = "black", weight = "bold", align = "center"),
    locations = cells_title("title")
  ) |> 
  tab_stubhead("Estimated Impact for Oklahomans with Misdemeanor Drug Charges") |> 
  gt_theme_nytimes()
  
```


## Possibly delete or comment out everything below

### Estimating December 2023
```{r}

## 7/9/2024 -- PR commenting this section out because returning an error when I try to render the HB3205 qmd file. Any changes made here were for that reason.


# ojodb <- ojo_connect()
# 
# start <- ymd("2010-01-01")
# end <- ymd("2023-12-01")
# 
# all_cases_big <- ojo_tbl("case", .con = ojodb) |>
#   filter(
#     case_type == "CM",
#     date_filed >= start,
#     date_filed < end
#   ) |>
#   select(
#     id,
#     case_type,
#     district,
#     date_filed,
#     counts,
#     open_counts
#   ) |>
#   mutate(
#     count = unnest(counts),
#     open_count = unnest(open_counts)
#   ) |>
#   left_join(
#     ojo_tbl("count", .con = ojodb),
#     by = c("count" = "id"),
#     suffix = c("", ".count")
#   ) |>
#   mutate(
#     count_as_filed = if_else(
#       is.na(open_count),
#       count_as_filed,
#       open_count
#     )
#   ) |>
#   collect()
# 
# # Defendants for all cases
# total_persons_big <- ojo_tbl("party", .con = ojodb) |>
#   right_join(
#     all_cases_big |>
#       distinct(id),
#     by = c("case_id" = "id"),
#     copy = TRUE
#   ) |>
#   filter(role == "Defendant") |>
#   distinct(oscn_id) |>
#   count() |>
#   collect()

# Pulling every case with at least one misdemeanor drug charge
# (excluding distribution/intent to dispense/etc.)
# case_misdemeanor_drug_big <- 
#   all_cases_big |>
#   filter(
#     str_detect(count_as_filed, drugs)
#   ) |>
#   filter(
#     !str_detect(count_as_filed, drug_exclude)
#   )

# Defendants in cases above:
# drug_misdemeanor_persons_big <- ojo_tbl("party", .con = ojodb) |>
#   right_join(
#     case_misdemeanor_drug_big |>
#       distinct(id),
#     by = c("case_id" = "id"),
#     copy = TRUE
#   ) |>
#   filter(role == "Defendant") |>
#   distinct(oscn_id) |>
#   count() |>
#   collect()

# A conservative estimate of the number of impacted individuals can be made by
# filtering misdemeanors by DPS violation code and related codes.
# For the most part, this excludes drug offenses occurring outside of motor vehicle
# case_dps_violation_big <- 
#   all_cases_big |>
#   filter(str_detect(count_as_filed, dps_codes),
#          !str_detect(count_as_filed, dps_exclude))

# This calculates number of defendants:
# dps_violation_persons_big <- ojo_tbl("party", .con = ojodb) |>
#   right_join(
#     case_dps_violation_big |>
#       distinct(id),
#     by = c("case_id" = "id"),
#     copy = TRUE
#   ) |>
#   filter(role == "Defendant") |>
#   distinct(oscn_id) |> 
#   count() |>
#   collect()
# 
# # Upper Bound including people from DPS OR misdemeanor method:
# both_methods_persons_big <- ojo_tbl("party", .con = ojodb) |>
#   right_join(
#     rbind(
#       case_dps_violation_big |>
#         select(id, date_filed),
#       case_misdemeanor_drug_big |>
#         select(id, date_filed)
#     ) |>
#       distinct(),
#     by = c("case_id" = "id"),
#     copy = TRUE
#   ) |>
#   filter(role == "Defendant") |>
#   distinct(oscn_id, .keep_all = TRUE) |> 
#   #distinct(oscn_id) |>
#   #count() |>
#   collect()

# Sanity check
# party table data starts in May 2021 so we actually only have Dec.2021 and Dec.2022 to estimate/impute the number of people charged in Dec.2023.
# ojo_tbl("party") |> 
#   select(created_at) |> 
#   mutate(year = floor_date(as_date(created_at), unit = "year")
#          ) |> 
#   count(year)

###################################################################################################
# Techniques to estimate the lower and upper bounds of individuals facing these charges in Dec 2023

# !!! need to use date_filed and not created_at
# note that party information was not tracked until 2021 and I'm assuming was back filled

# this looks at 2022-01-01 through 2023-12-01
# dec_2022 <- both_methods_persons |> 
#   mutate(month_year = floor_date(as_date(date_filed), unit = "month")
#   ) |> 
#   count(month_year) |> 
#   filter(month_year == "2022-12-01")

# this captures 2010-01-01 through 2023-12-01
# the monthly counts slightly differ depending on which chunk of time we're looking at
# all_dec <- both_methods_persons_big |> 
#   mutate(month_year = 
#            floor_date(as_date(date_filed), unit = "month"), 
#          month = month(as_date(date_filed))
#          ) |> 
#   group_by(month_year, month) |> 
#   count(month_year) |> 
#   filter(month == 12)

# option 1) Just use Dec 2022 -- kind of dumb but the easiest route
# both_methods_persons |> 
#   distinct(oscn_id) |>
#   count() |> 
#   rename(upper = n) |> 
#   mutate(upper = upper + dec_2022$n) |> 
#   bind_cols(
#   dps_violation_persons) |> 
#   rename(lower = n)
# 


```

