---
title: HB 1795 Impact Estimate
subtitle: Open Justice Oklahoma
format:
  ojo-report-template-html: default
author: Polina Rozhkova
date: last-modified
logo: "www/ojo-logo-white.png"
number-sections: true
smooth-scroll: true
title-block-banner: "black"
title-block-banner-color: "white"
knitr:
  opts_chunk:
    fig.align: center
    echo: FALSE
    message: FALSE
    warning: FALSE
description: Estimating the number of Oklahomans impacted by the amended provision concerning driver license revocations for individuals with misdemeanor drug charges. Analysis conducted at the request of Oklahomans for Criminal Justice Reform (OCJR) and ProsperOK.
abstract-title: "Abstract"
abstract: Between 8107 and 13485 Oklahomans with misdemeanor drug charges were impacted by provisions in HB 1795 between January 1, 2022 and June 30, 2023.

---
```{r}
library(readr)
library(stringr)
library(readxl)
library(tidycensus)
library(tidyverse)
library(here)
library(janitor)
library(glue)
library(ojoverse)
library(lubridate)
library(gt)
library(downloadthis)
```

## Background

Oklahoma’s House Bill 1795 went into effect November 1, 2021 (Q2 FY 2022). This new law reduces the number of offenses that can lead to a license revocation, although certain circumstances may still result in revocation. Additionally, HB 1795 aims to make provisional licenses more accessible and affordable. The bill introduces restrictions on license suspensions due to failure to pay, and it included an important provision that reduces the minimum payment plan for outstanding fines and fees. Previously set at \$25 per month, the HB 1795 lowers the monthly payment to \$5. The aforementioned fees encompass various charges such as provisional license fees, warrant fees, court costs or fees, and driver license reinstatement fees. 

## Data and Methodology

This analysis uses Oklahoma court records sourced by OSCN and ODCR beginning January 1, 2022 (Q3 FY 2022) to estimate the impact of HB 1795 during a period of full implementation. We explore two provisions of the bill in order to estimate lower and upper bounds for the number of individuals with misdemeanor drug charges that were impacted by the bill’s passing. For the lower bound, we focus on the provision concerned with modifying the types of offenses requiring immediate revocation of driving privileges (Section 6-205). Specifically we account for individuals who are charged with possession of a controlled dangerous substance as defined by 

A conservative estimate of the number of impacted individuals is made by filtering drug-related misdemeanors by DPS violation code and similar codes.

For an estimate of the upper bound, we account for individuals who may be subject to fines and fees as a result 


## Results

```{r}
# ojo_tbl("party") |>
#   right_join(
#     rbind(case_dps_violation |> select(id), case_misdemeanor_drug |> select(id)) |> distinct(),
#     by = c("case_id" = "id"),
#     copy = TRUE
#   ) |>
#   filter(role == "Defendant") |>
#   distinct(oscn_id) |>
#   count() |> 
#   gt()
```
